# Basic tour of the R language
## Variable assignment and function invocation
 In R, we can create variables by simply assigning them a value.

```{r}
    x <- 3
    x 
```

 We don't have to declare them with a special keyword, nor do we have
 to give them a type.

 If we try and access a variable that doesn't exist, we are quickly
 told that we are going down the wrong path!
 
 ```{r}
 z
 ```

 Note that the assignment is via an arrow. This is a nice peculiar
 feature of R.

 Actually a regular old = sign does work, apart from some strange cases
 where it does not. These reasosn make sense when you start to
 understand some more advanced parts of the language, or so I am told.

 To be honest, I generally go for the = sign, but I know that I have
 been hurt by this a few times so it really is best to stick with the arrows.

Similar to other langauges, we can call and use functions. R has a lot
of useful built in functions for dealing with numeric data and
vectors.

This one produces a sequence of numbers up to the value you give it.

 The standard with R seems to be that there are shorthand versions of
 the functions but if you want to give it more or different parameters,
 you have to give them names like this.

```{r}
    x <- seq(10)
    x <- seq(from=0,by=10,length=10)
```

 There are a whole heap of other standard functions, like rnorm
 here that will give us 10 values from a normal
 distribution. Unsuprisingly, regular arithmetic operators are also available!

```{r}
    x <- rnorm(10)
    x <- 3
    y <- x+2
```

## Data types

An important concept to understand when learning R is that although these look like
regular numbers or integers, our basic core type is actually a vector.

```{r}
    x <- 3
    len(x)
```

So here, although it looks like X is 3, it's really better to think
about x being a vector of length 1.

We can create longer vectors with the c 'combine' function.  You maybe
could see this as similar to array construction with angle braces in
other languages like ruby.


```{r}
    x <- c(1,2,3,4,5)

    x
```

This might seem strange, but it's kind of cool because it means that
any function or operation with which works on 1 value is generally
able to work on N values.

For example check out the result of the following operation from
before when x is a vector... it's another vector.

```{r}
    y <- x+3
    y
```

What this means is that you can start to make code that looks
an awful lot like the symbols we might use on paper perform as
it should, like quadratic equations we probably all covered at school.

```{r}
    
    x <- seq(100)
    
    y <- x^2 + 4
```

Of course there are several other fundamental data types,
including the more sophisticated relation of the vector -
matrices and lists that behave a little like hashes/dictionaries.

```{r}
   
    m <- matrix(c(1,2,3,4,5,6), nrow=2)
    m

    myList <- list(x=2, y=2)
    myList$x
    
```

The at the more fundamental level, we have numeric data,
characters and things called 'factors'. Which are similar to
enumerations. These can come in pretty handy as you might see
later for categorical data.

```{r} 
x <- "character data"
x
x <- 2.2
x
factor(c('Red', 'Green', 'Blue'))
```
## Plotting

And of course, it wouldn't be stats without any kind of plotting. R has some simple, but powerful basic plotting support built in.

The basic plot() function covers a lot of bases.
```{r}
 

    x <- seq(from=-1,to=1,length=1000)
    
    y <- x^2 + 4
    
    plot(x,y)

```

See also ```barplot```, ```hist```, ```boxplot```

 We can tweak the parameters a little bit - the documentation covers
 this in more detail.
 
```{r}

    plot(x,y,type='l',col='green')

    
    n <- rnorm(100)
    plot(n)
    barplot(n)
    hist(n)
    boxplot(n)
```


 I got this one from the internet....

```{r}
  
    t <- seq(from=-10,to=10, length=1000)
    x <- sin(t)^3
    y <- 13 * cos(t) - 5* (cos(2 * t)) - cos(4 * t)
    plot(x,y, type='l')
    
    polygon(x,y,col='red')
```

## Data frames

Before getting into the real demo, the last thing I wanted to show you
was the data.frame object.

Rather than having lots of different data variables like my previous
examples, I can group together lots of variables in one data frame.


```{r}
    my.data <- data.frame(col=rep(c('red','green'),times=120), size=rpois(120, 100))
    my.data$col
    
    #View(my.data)
    
    my.data$shape <- rep(c('triangle', 'circle', 'square'))
```

I can subset a data frame - filter down to rows with particular attributes:

Later on you will see what we can do with these, particularly when we use these to hold external data.

```{r}
    red.triangles <- my.data[my.data$shape == 'circle' && my.data$col == 'red',]
    
    summary(my.data)
```